---
title: "CMU 2022 Football Scrape"
author: "Jordan Gilbert, Marion Haney"
date: "2/1/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rvest)
require(magrittr)
require(stringr)
require(purrr)
require(dplyr)
```


Function to get urls for a number of weeks for a given year
```{r}
# Default: 2022 regular season games
seasonSchedule = function(year = 2022, weeks = 10) {
  # check that valid number of games
  weeks = ifelse(weeks > 12, 10, weeks)
  weeks = seq(weeks)
  seasonLink = c()
  for (week in weeks) {
    weekLink = paste(c("https://www.d3football.com/scoreboard/", year, 
                     "/composite?view=", week), collapse = "")
    seasonLink = c(seasonLink, weekLink)
  }
  return(seasonLink)
}
URLs = seasonSchedule(year = 2022, weeks = 10)
```



Function to get play by play data:
```{r}
gamePlayByPlay = function(game, date, time, location, homeTeam, homeAbb, awayTeam, awayAbb) {
  downs = c("1", "2", "3", "4", "OT")
  quarter = list()
  possession = list()
  down = list()
  distancetoGo = list()
  distancetoGoal = list()
  playType = list()
  netYards = list()
  scoreHome = list()
  scoreAway = list()
  Outcome = list()
  numDrives = list()
  totalPlays = list()
  binaryPointsScored = list()
  binaryCompletedPass = list()
  binaryIncompletePass = list()
  binaryRun = list()
  binaryTouchdown = list()
  binaryFieldGoal = list()
  binaryPenalty = list()
  binaryPunt = list()
  binarySack = list()
  binaryInterception = list()
  binarySafety = list()
  binaryFumble = list()
  binaryExtraPoint = list()
  currentQuarter = NULL
  lastPossession = "NULL"
  currentPossession = NULL
  driveNumberAway = 0
  driveNumberHome = 0
  driveNumber = 0
  numberOfPlaysAway = 0
  numberOfPlaysHome = 0
  numPlays = 0
  
  for (i in 1:nrow(game)) {
    # Binary Outcomes
    is.pointsScored = FALSE
    is.completedPass = FALSE
    is.incompletePass = FALSE
    is.run = FALSE
    is.touchdown = FALSE
    is.fieldGoal = FALSE
    is.penalty = FALSE
    is.punt = FALSE
    is.sack = FALSE
    is.intercepted = FALSE
    is.safety = FALSE
    is.fumble = FALSE
    is.extraPoint = FALSE
    
    # Information about Game
    situation = game$Situation[i]
    if (situation == "") {
      situation = game$Situation[i-1]
    }
    play = game$Play[i]
    startsWith = substr(situation, 1, 1)
    
    # Gets Quarter of the Game
    if (nchar(situation) <= 3 && startsWith %in% downs) {
      currentQuarter = situation
      # no other information, so skip
      next
    }
    
    # Gets Possession at Time
    if (str_detect(situation, "[0-1][0-9]:[0-5][0-9]") & !str_detect(situation, "elapsed")) {
      currentPossession = sub(" at.*", "", situation)
      # update drive number and possession
      if (lastPossession != currentPossession) {
        if (currentPossession == homeTeam) {
          driveNumberHome = driveNumberHome + 1
          driveNumber = driveNumberHome
        }
        else {
          driveNumberAway = driveNumberAway + 1
          driveNumber = driveNumberAway
        }
        lastPossession = currentPossession
      }
      # no other information, so skip
      next
    }
    
    # Looks at columns with plays only
    if (!(grepl("[1-4][(st)|(nd)|(rd)|(th)]", substr(situation, 1, 3)))) {
      # has no wanted information, so skip
      next
    }
    if (situation == play) {
      # has no wanted information, so skip
      next
    }
    
    # If just update on score, skip
    if (grepl("([[:alpha:]]|[[:punct:]]|[[:space:]])+ [[:digit:]]+,
              ([[:alpha:]]|[[:punct:]]|[[:space:]])+ [[:digit:]]", play)) {
      # has no wanted information, so skip
      next
    }
    
    # if recap of field position, skip
    if (grepl("[1-4][[:alpha:]]{2} and [[:digit:]]+.", play)) {
      # has no wanted information, so skip
      next
    }
    
    # if not valid play, skip
    if (nchar(play) <= 1) {
      # has no wanted information, so skip
      next
    }
    
    # Plays with information
    if (nchar(situation) >= 25 && startsWith %in% downs) {
      
      # down
      currentDown = startsWith
      # distance to first down
      currentDistanceToGo = gsub(".*and (.+) at.*", "\\1", situation)
      
      if (!(is.na(str_extract(play,
          "(kickoff|hurry|ball|drive|quarter|half|Timeout|game|toss|QB|receive)")))) {
        # has no wanted information, so skip
        next
      }
      
      # abbreviation of location on field
      abb = gsub(".*at (.+)", "\\1", situation)
      lets = unlist(str_extract_all(abb, "[[:alpha:]]"))
      toGoal = (unlist(str_extract_all(abb, "[[:digit:]]")))
      abb = str_c(lets, collapse = "")
      toGoal = as.numeric(str_c(toGoal, collapse = ""))
      
      # get distance from goal
      if (currentPossession == homeTeam | str_detect(str_to_lower(currentPossession),
              str_to_lower(homeTeam)) |
          str_detect(str_to_lower(homeTeam), str_to_lower(currentPossession))) {
        if (abb == homeAbb) {
          currentDistanceToGoal = as.character(100 - toGoal)
        }
        else {
          currentDistanceToGoal = as.character(toGoal)
        }
      }
      else {
        if (abb == awayAbb) {
          currentDistanceToGoal = as.character(100 - toGoal)
        }
        else {
          currentDistanceToGoal = as.character(toGoal)
        }
      }
      
      # get play
      currentPlay = str_extract(play, 
          "(rush|pass|PENALTY|fumble|sacked|punt|field goal|kick|Kneel)")
      
      if (is.na(currentPlay)) {
        #print(play)
        #print(game$Play[i-1])
        yards = "Error"
        currentPlay = "NA"
      }
      
      if (str_detect(play, "PENALTY")) {
        yards = "Penalty"
        is.penalty = TRUE
      }
      
      if (str_detect(play, "flea-flicker")) {
        yards = "Error"
      }
      
      else if (currentPlay == "kick") {
        yards = "Extra Point Attempt"
      }
      
      else if (currentPlay == "rush") {
        if (!(is.na(str_extract(play, "no gain")))) {
          yards = "0"
        }
        else {
          yards = (sub(".*for (.+) yard.*", "\\1", play))
          if (!(is.na(str_extract(yards, "loss")))) {
            yards = gsub(".*of (.+)", "\\1", yards)
            yards = as.character(as.numeric(yards) * -1)
          }
        }
        is.run = TRUE
      }
      else if (currentPlay == "pass") {
        if (!(is.na(str_extract(play, "incomplete")))) {
          yards = "0"
          is.incompletePass = TRUE
        }
      else if (!(is.na(str_extract(play, "intercepted")))) {
          yards = "Intercepted"
          is.incompletePass = TRUE
          is.intercepted = TRUE
        }
      else {
          yards = (sub(".*for (.+) yard.*", "\\1", play))
          if (!(is.na(str_extract(yards, "loss")))) {
            yards = gsub(".*of (.+)", "\\1", yards)
            yards = as.character(as.numeric(yards) * -1)
          }
          is.completedPass = TRUE
        }
      }
        
      else if (currentPlay == "Kneel") {
        if (!(is.na(str_extract(play, "no gain")))) {
          yards = "0"
        }
        else {
          yards = (sub(".*for (.+) yard.*", "\\1", play))
          if (!(is.na(str_extract(yards, "loss")))) {
            yards = gsub(".*of (.+)", "\\1", yards)
            yards = as.character(as.numeric(yards) * -1)
          }
        }
      }
      else if (currentPlay == "fumble") {
        yards = "Fumble"
        is.fumble = TRUE
      }
      else if (currentPlay == "punt") {
        yards = "Punt"
        is.punt = TRUE
      }
      else if (currentPlay == "field goal") {
        yards = "Field Goal"
        is.fieldGoal = TRUE
      }
      else if (currentPlay == "sacked") {
        yards = (sub(".*for (.+) yard.*", "\\1", play))
        if (!(is.na(str_extract(yards, "loss")))) {
          yards = gsub(".*of (.+)", "\\1", yards)
          yards = as.character(as.numeric(yards) * -1)
        }
        is.sack = TRUE
      }
      
      # get outcome
      play = tolower(play)
      touch = !(is.na(str_extract(play, 
      "(touchdown|penalty|fumble|sacked|punt|incomplete|field goal|intercepted|kick|safety)")))
      if (touch) {
        touch = str_extract_all(play, 
      "(touchdown|penalty|fumble|sacked|punt|incomplete|field goal|intercepted|kick|safety)")
        if (touch == "touchdown") {
          outcome = "Touchdown"
          is.touchdown = TRUE
          is.pointsScored = TRUE
        }
        else if (touch == "penalty") {
          outcome = "Penalty"
          is.penalty = TRUE
        }
        else if (touch == "fumble") {
          outcome = "Fumble"
          is.fumble = TRUE
        }
        else if (touch == "sacked") {
          outcome = "Sack"
          is.sack = TRUE
        }
        else if (touch == "punt") {
          outcome = "Punt"
          is.punt = TRUE
        }
        else if (touch == "incomplete") {
          outcome = "Incomplete Pass"
          is.incompletePass = TRUE
        }
        else if (touch == "field goal") {
          if (!(is.na(str_extract(play, "missed")))) {
            outcome = "Missed Field Goal"         
          }
          else {
            outcome = "Field Goal"
            is.pointsScored = TRUE
          }
          is.fieldGoal = TRUE
        }
        else if (touch == "intercepted") {
          outcome = "Interception"
          is.intercepted = TRUE
        }
        else if (touch == "kick") {
          if (!(is.na(str_extract(play, "missed")))) {
            outcome = "Missed Extra Point Attempt"
          }
          else {
            outcome = "Made Extra Point Attempt"
            is.pointsScored = TRUE
          }
          is.extraPoint = TRUE
        }
        else if (touch == "safety") {
          outcome = "Safety"
          is.safety = TRUE
        }
        else {
          outcome = str_to_lower(paste(unlist(touch), collapse = ", "))
          if (str_detect(outcome, "fumble")) {
            is.fumble = TRUE
          }
          if (str_detect(outcome, "penalty")) {
            is.penalty = TRUE
          }
          if (str_detect(outcome, "safety")) {
            is.safety = TRUE
          }
          if (str_detect(outcome, "kick")) {
            is.extraPoint = TRUE
            if (str_detect(play, "attempt good")) {
              is.pointsScored = TRUE
            }
          }
        }
      }
      else {
        outcome = ""
      }
      
      # total number of plays
      if (currentPossession == homeTeam) {
        if (!(currentPlay == "PENALTY" | str_detect(play, "NO PLAY")) | 
            str_detect(play, "flea-flicker")) {
          numberOfPlaysHome = numberOfPlaysHome + 1
          numPlays = numberOfPlaysHome
        }
      }
      else {
        if (!(currentPlay == "PENALTY" | str_detect(play, "NO PLAY")) | 
            str_detect(play, "flea-flicker")) {
          numberOfPlaysAway = numberOfPlaysAway + 1
          numPlays = numberOfPlaysAway
        }
      }
    }
    
    quarter = c(quarter, list(currentQuarter))
    possession = c(possession, list(currentPossession))
    down = c(down, list(currentDown))
    distancetoGo = c(distancetoGo, list(currentDistanceToGo))
    playType = c(playType, list(currentPlay))
    netYards = c(netYards, list(yards))
    Outcome = c(Outcome, list(outcome))
    distancetoGoal = c(distancetoGoal, list(currentDistanceToGoal))
    numDrives = c(numDrives, list(driveNumber))
    totalPlays = c(totalPlays, list(numPlays))
    binaryPointsScored = c(binaryPointsScored, list(is.pointsScored))
    binaryCompletedPass = c(binaryCompletedPass, list(is.completedPass))
    binaryIncompletePass = c(binaryIncompletePass, list(is.incompletePass))
    binaryRun = c(binaryRun, list(is.run))
    binaryTouchdown = c(binaryTouchdown, list(is.touchdown))
    binaryFieldGoal = c(binaryFieldGoal, list(is.fieldGoal))
    binaryPenalty = c(binaryPenalty, list(is.penalty))
    binaryPunt = c(binaryPunt, list(is.punt))
    binarySack = c(binarySack, list(is.sack))
    binaryInterception = c(binaryInterception, list(is.intercepted))
    binarySafety = c(binarySafety, list(is.safety))
    binaryFumble = c(binaryFumble, list(is.fumble))
    binaryExtraPoint = c(binaryExtraPoint, list(is.extraPoint))
  }
  
  PlaybyPlayTable = data.frame(Date = date, Time = time, Location = location, 
                               "Home Team" = homeTeam, 
                               "Home Team Abbreviation" = homeAbb, 
                               "Away Team" = awayTeam, 
                               "Away Team Abbreviation" = awayAbb, 
                               Quarter = unlist(quarter), 
                               "Team Possession" = unlist(possession),
                               Down = unlist(down), 
                               "Distance to First Down" = unlist(distancetoGo), 
                               "Distance to Touchdown" = unlist(distancetoGoal), 
                               "Play Type" = unlist(playType), 
                               "Yards" = unlist(netYards), Outcome = unlist(Outcome), 
                               "Points Scored?" = unlist(binaryPointsScored), 
                               "Completed Pass?" = unlist(binaryCompletedPass), 
                               "Incomplete Pass?" = unlist(binaryIncompletePass),
                               "Run?" = unlist(binaryRun), 
                               "Touchdown?" = unlist(binaryTouchdown), 
                               "Field Goal?" = unlist(binaryFieldGoal), 
                               "Penalty?" = unlist(binaryPenalty), 
                               "Punt?" = unlist(binaryPunt), 
                               "Sack?" = unlist(binaryPunt), 
                               "Interception?" = unlist(binaryInterception), 
                               "Safety?" = unlist(binarySafety), 
                               "Fumble?" = unlist(binaryFumble), 
                               "Extra Point?" = unlist(binaryExtraPoint), 
                               "Drive Number" = unlist(numDrives), 
                               "Team Play Number" = unlist(totalPlays))
  
  return(PlaybyPlayTable)
}
```


Function to get urls of play by plays from a given week
```{r}
weekSchedule = function(url) {
  # read schedule of given week
  schedule = read_html(url, as.data.frame = TRUE, stringsAsFactors = TRUE)
  x = schedule %>% html_nodes("table") %>% .[[1]]
  
  # schedule as table
  fullSchedule = x %>% html_table(fill = TRUE)
  
  # all links on table
  allLinks = x %>% html_nodes("a") %>% html_attr("href") %>% grep("boxscores", ., value = TRUE)
  
  # links to play by play
  links = unlist(lapply(allLinks, function(x) {begin = "https://www.d3football.com"; 
          end = "?view=plays"; adding = paste(begin, x, sep = ""); 
          ending = paste(adding, end, sep = "");
          return(ending)}))
}
```


Function to gather multiple weeks of data for one team, given the URLs and team name
```{r}
# combining the weeks together

# function for combining play-by-play data for several links
# input: listOfUrls list of links, string team
# output: data frame of play-by-play data for a given team

combineSeasonData = function(listOfUrls, teamName) {
  for (m in 1:length(listOfUrls)) {
      url = listOfUrls[m]
      links = weekSchedule(url)
      
      bigBadTable = NULL
      for (series in links) {
        # play by play of game
        gameSeries = read_html(series, as.data.frame = TRUE, stringsAsFactors = TRUE)
        y = gameSeries %>% html_nodes("table") %>% .[[3]]
        
        # play by play as table
        game = y %>% html_table(fill = TRUE)
        colnames(game) = c("Situation", "Play")
        
        # getting teams
        awayInfo = gameSeries %>% html_nodes("td") %>% .[[2]] %>% html_text()
        awayInfo = unlist(str_extract_all(awayInfo, "[^\n\r\t]"))
        awayInfo = str_c(awayInfo, collapse = "")
        awayInfo = unlist(str_split(awayInfo, "   +"))
        awayTeam = awayInfo[2]
        awayTeamScore = awayInfo[3]
        homeInfo = gameSeries %>% html_nodes("td") %>% .[[3]] %>% html_text()
        homeInfo = unlist(str_extract_all(homeInfo, "[^\n\r\t]"))
        homeInfo = str_c(homeInfo, collapse = "")
        homeInfo = unlist(str_split(homeInfo, "   +"))
        homeTeam = homeInfo[2]
        homeTeamScore = homeInfo[3]
        # only certain teams
        if (homeTeam != teamName && awayTeam != teamName) { next }
        
        # getting teams, location, date, time
        setting = gameSeries %>% html_nodes("td") %>% .[[1]] %>% html_text()
        setting = unlist(str_extract_all(setting, "[^\n\r\t]"))
        setting = str_c(setting, collapse = "")
        setting = unlist(str_split(setting, "  "))
        teamsLocation = setting[2]
        teamsLocation = unlist(str_split(teamsLocation, " - "))
        location = teamsLocation[2]
        dateTime = setting[3]
        dateTime = unlist(str_split(dateTime, " - "))
        date = dateTime[1]
        time = dateTime[2]
        
        awayAbb = gsub(".*at (.+)", "\\1", game[3, ]$Situation)
        awayAbb = unlist(str_extract_all(awayAbb, "[[:alpha:]]"))
        awayAbb = str_c(awayAbb, collapse = "")
        homeAbbFound = FALSE
        index = 4
        i = 1
        
        while (!homeAbbFound) {
          startOfGame = game[i,]
          if (grepl("[1-4][(st)|(nd)|(rd)|(th)]", substr(startOfGame$Situation, 1, 3))) {
            homeAbb = gsub(".*at (.+)", "\\1", startOfGame$Situation)
            homeAbb = unlist(str_extract_all(homeAbb, "[[:alpha:]]"))
            homeAbb = str_c(homeAbb, collapse = "")
            if (homeAbb != awayAbb) {
              homeAbbFound = TRUE
            }
          }
          i = i + 1
        }
        
        playByPlay = gamePlayByPlay(game, date, time, location, homeTeam, homeAbb, awayTeam, awayAbb)
        if (m == 1) {
          df = data.frame(playByPlay)
        } else {
          bigBadTable = playByPlay
          df = rbind(df, bigBadTable)
        }
      }
    }
    return(df)
}

```


Function to batch scrape multiple teams given urls of the season and week
```{r}
batchScrape = function(urls, teams) {
  df = data.frame()
  for (i in 1:length(teams)) {
    teamName = teams[i]
    print(teamName)
    newDf = combineSeasonData(urls, teamName)
    df = rbind(df, newDf)
  }
  df2 = df[!duplicated(df), ] # remove duplicate rows
  return(df2)
}
```


```{r}
batchScrape2 = function(urls, teams) {
  df = data.frame()
  for (i in 1:length(teams)) {
    teamName = teams[i]
    print(teamName)
    newDf = NULL
    
    # make newDf
    for (m in 1:length(urls)) {
      url = urls[m]
      links = weekSchedule(url) # links to the play by plays
      
      bigBadTable = NULL
      for (series in links) {
        # play by play of game
        gameSeries = read_html(series, as.data.frame = TRUE, stringsAsFactors = TRUE)
        y = gameSeries %>% html_nodes("table") %>% .[[3]]
        
        # play by play as table
        game = y %>% html_table(fill = TRUE)
        colnames(game) = c("Situation", "Play")
        
        # getting teams
        awayInfo = gameSeries %>% html_nodes("td") %>% .[[2]] %>% html_text()
        awayInfo = unlist(str_extract_all(awayInfo, "[^\n\r\t]"))
        awayInfo = str_c(awayInfo, collapse = "")
        awayInfo = unlist(str_split(awayInfo, "   +"))
        awayTeam = awayInfo[2]
        awayTeamScore = awayInfo[3]
        homeInfo = gameSeries %>% html_nodes("td") %>% .[[3]] %>% html_text()
        homeInfo = unlist(str_extract_all(homeInfo, "[^\n\r\t]"))
        homeInfo = str_c(homeInfo, collapse = "")
        homeInfo = unlist(str_split(homeInfo, "   +"))
        homeTeam = homeInfo[2]
        homeTeamScore = homeInfo[3]
        # only certain teams
        if (homeTeam != teamName && awayTeam != teamName) { next }
        
        # getting teams, location, date, time
        setting = gameSeries %>% html_nodes("td") %>% .[[1]] %>% html_text()
        setting = unlist(str_extract_all(setting, "[^\n\r\t]"))
        setting = str_c(setting, collapse = "")
        setting = unlist(str_split(setting, "  "))
        teamsLocation = setting[2]
        teamsLocation = unlist(str_split(teamsLocation, " - "))
        location = teamsLocation[2]
        dateTime = setting[3]
        dateTime = unlist(str_split(dateTime, " - "))
        date = dateTime[1]
        time = dateTime[2]
        
        awayAbb = gsub(".*at (.+)", "\\1", game[3, ]$Situation)
        awayAbb = unlist(str_extract_all(awayAbb, "[[:alpha:]]"))
        awayAbb = str_c(awayAbb, collapse = "")
        homeAbbFound = FALSE
        index = 4
        i = 1
        
        while (!homeAbbFound) {
          startOfGame = game[i,]
          if (grepl("[1-4][(st)|(nd)|(rd)|(th)]", substr(startOfGame$Situation, 1, 3))) {
            homeAbb = gsub(".*at (.+)", "\\1", startOfGame$Situation)
            homeAbb = unlist(str_extract_all(homeAbb, "[[:alpha:]]"))
            homeAbb = str_c(homeAbb, collapse = "")
            if (homeAbb != awayAbb) {
              homeAbbFound = TRUE
            }
          }
          i = i + 1
        }
        
        playByPlay = gamePlayByPlay(game, date, time, location, homeTeam, homeAbb, awayTeam, awayAbb)
        if (m == 1) {
          newDf = data.frame(playByPlay)
        } else {
          bigBadTable = playByPlay
          newDf = rbind(newDf, bigBadTable)
        }
      }
    }
  }
    
    df = rbind(df, newDf)
  
  
  
  df2 = distinct(df) # remove duplicate rows
  return(df2)
}

```


Test:
Get URLs & teams
```{r}
PAC = c("Allegheny", "Bethany", "Carnegie Mellon", "Case Western Reserve", "Geneva",
        "Grove City", "St. Vincent", "Thiel", "Washington and Jefferson", 
        "Waynesburg", "Westminster (Pa.)")
```


Batch Scrape from URLs
```{r}

# make df
checkURLs = c("https://www.d3football.com/scoreboard/2022/composite?view=1", "https://www.d3football.com/scoreboard/2022/composite?view=2")

df = batchScrape(checkURLs, PAC)
```


Save data frame to local
```{r}
# save df
#write.csv(df, "CMU-2022-pbp.csv", row.names=FALSE)
```


