---
title: "Modeling-Marion"
author: "Marion Haney"
date: "2023-04-10"
output: html_document
---

Load in libraries
```{r}
# Libraries
require(ggplot2)
require(tidyverse)
require(dplyr)
require(nnet)
require(nnet)
require(caret)
```

Load in data
```{r}
df <- read.csv("PAC-2022-Cleaned-pbp2.csv")
cmu.df <- read.csv("PAC-2022-Cleaned-pbp2.csv") # to subset later
```

Reorder the factors of Next_Score_Half
```{r}
df <- df[df$Next_Score_Half %in% c("Field_Goal", "Opp_Field_Goal", "Touchdown", "Opp_Touchdown", "No_Score"),]
df$Next_Score_Half <- factor(df$Next_Score_Half, levels = c("No_Score", 
                                                            "Field_Goal", 
                                                            "Opp_Field_Goal",
                                                            "Touchdown",
                                                            "Opp_Touchdown"))
```


Cleaning Distance.to.First.Down
```{r}
df$Distance.to.First.Down2 <- as.numeric(ifelse(df$Distance.to.First.Down == "Goal",
                                                df$Distance.to.Touchdown,
                                                df$Distance.to.First.Down))
df$is.Goal.Down <- ifelse(df$Distance.to.First.Down == "Goal",
                          1, 0)
```

Add a game.ID column and a fold column
```{r}
df$game.ID <- paste(df$Date, df$Home.Team.Abbreviation, "x", df$Away.Team.Abbreviation, sep="")

# fold
set.seed(123)
rand_vals <- sample(1:5, nrow(df), replace=TRUE)
df$fold <- rand_vals
```


Multinomial logistic regression for next score using down and distance (yds) to touchdown
```{r}
ep.model <- multinom(factor(Next_Score_Half) ~ 
                             factor(Down)
                           + log(Distance.to.First.Down2)
                           + Distance.to.Touchdown
                           + factor(is.Goal.Down)
                           + factor(Down):log(Distance.to.First.Down2), 
                           data = df, maxit = 300)
summary(ep.model)
```

Model Visualizations:
From: 
https://www.stat.cmu.edu/cmsac/sure/2022/materials/lectures/slides/21-Multinomial-multilevel.html#20 
*** Change to CV by fold/fold_ids
```{r}
init_loso_cv_preds <- df %>%
  group_by(fold) %>%
  mutate(train = row_number() <= round(0.75 * n())) %>%
  ungroup() %>%
  group_split() %>%
  map(function(split) {
    # Separate test and training data:
    train_data <- split %>% filter(train)
    test_data <- split %>% filter(!train)

    # Fit multinomial logistic regression model:
    ep_model <- multinom(factor(Next_Score_Half) ~ 
                             factor(Down)
                           + log(Distance.to.First.Down2)
                           + Distance.to.Touchdown
                           + factor(is.Goal.Down)
                           + factor(Down):log(Distance.to.First.Down2), 
                           data = df, maxit = 300)

    # Return dataset of class probabilities:
    predict(ep_model, newdata = test_data, type = "probs") %>%
      as_tibble() %>%
      mutate(Next_Score_Half = test_data$Next_Score_Half)
  }) %>%
  bind_rows()
```

Calibration results for each scoring event
```{r}
ep_cv_loso_calibration_results <- init_loso_cv_preds %>%
  pivot_longer(No_Score:Opp_Touchdown,
               names_to = "next_score_type",
               values_to = "pred_prob") %>%
  mutate(bin_pred_prob = round(pred_prob / 0.05) * .05) %>%
  group_by(next_score_type, bin_pred_prob) %>%
  summarize(n_plays = n(), 
            n_scoring_event = length(which(Next_Score_Half == next_score_type)),
            bin_actual_prob = n_scoring_event / n_plays,
            bin_se = sqrt((bin_actual_prob * (1 - bin_actual_prob)) / n_plays)) %>%
  ungroup() %>%
  mutate(bin_upper = pmin(bin_actual_prob + 2 * bin_se, 1),
         bin_lower = pmax(bin_actual_prob - 2 * bin_se, 0))
```

```{r}
ep_cv_loso_calibration_results <- na.omit(ep_cv_loso_calibration_results)
ep_cv_loso_calibration_results %>%
  mutate(next_score_type = fct_relevel(next_score_type, "No_Score", "Field_Goal", 
                                       "Opp_Field_Goal", "Touchdown", "Opp_Touchdown"),
  next_score_type = fct_recode(next_score_type,
                               "No Score (0)" = "No_Score", 
                               "Field Goal (3)" = "Field_Goal",
                               "Opponent Field Goal (-3)" = "Opp_Field_Goal",
                               "Touchdown (7)" = "Touchdown",
                               "Opponent Touchdown (-7)" = "Opp_Touchdown")) %>%
  ggplot(aes(x = bin_pred_prob, y = bin_actual_prob)) +
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = n_plays)) +
  geom_errorbar(aes(ymin = bin_lower, ymax = bin_upper)) + #coord_equal() +   
  scale_x_continuous(limits = c(0,1)) + 
  scale_y_continuous(limits = c(0,1)) + 
  labs(size = "Number of plays", x = "Estimated next score probability", 
       y = "Observed next score probability") + 
  theme_bw() + 
  ggtitle("Calibration for EP Model Predictions") +
  theme(strip.background = element_blank(), 
        axis.text.x = element_text(angle = 90), 
        legend.position = c(0.95, -0.05), legend.justification = c(1, 0)) +
  facet_wrap(~ next_score_type, ncol = 3)
```

Expected Points Column
```{r}
# get base probabilities
base_ep_preds <- as.data.frame(predict(ep.model, newdata = df, type = "probs"))
base_ep_preds <- rename(base_ep_preds,
                        No_Score_Prob=No_Score,
                        Field_Goal_Prob=Field_Goal,
                        Opp_Field_Goal_Prob=Opp_Field_Goal,
                        Touchdown_Prob=Touchdown,
                        Opp_Touchdown_Prob=Opp_Touchdown)
```

```{r}
# add to dftg
df <- cbind(df, base_ep_preds)
```

```{r}
EP <- rep(0,7787)
# multiply by point values
for (ii in 1:nrow(df)) {
  EP[[ii]] = sum(c(df[ii,43]*0, df[ii,44]*3, df[ii,45]*-3, df[ii,46]*7, df[ii,47]*-7))
}
df$EP <- EP
```


Expected Points Added Column
```{r}
df <- df %>%
  group_by(game.ID) %>%
  mutate(EPA =
           ifelse(Team.Possession == lead(Team.Possession),
                  lead(EP) - EP,
                  -lead(EP) - EP)) %>%
  ungroup()
```

Making EPA for scoring plays the score added - EP:
```{r}
for (i in 1:(nrow(df)-1)) {
  if (!df[[i, "Points.Scored."]]) next
  else {
    if (df[[i, "Touchdown."]]) {
      if (df[[i,"Turnover."]]) {
        # defensive touchdown
        df[i, "EPA"] = -7 - df[i, "EP"]
      } else {
        # offensive touchdown
        df[i, "EPA"] = 7 - df[i, "EP"]
      }
    } else if (df[[i, "Field.Goal."]]) {
      # offensive field goal
      df[i, "EPA"] = 3 - df[i, "EP"]
    }
  }
}
```


Save the EP/EPA columns to the df
```{r}
write.csv(df, "PAC-2022-cleaned-pbp-wEPA.csv")
```


## JUST CMU
Reorder the factors of Next_Score_Half
```{r}
cmu.df <- cmu.df[cmu.df$Next_Score_Half %in% c("Field_Goal", "Opp_Field_Goal", "Touchdown", "Opp_Touchdown", "No_Score"),]
cmu.df$Next_Score_Half <- factor(cmu.df$Next_Score_Half, levels = c("No_Score", 
                                                            "Field_Goal", 
                                                            "Opp_Field_Goal",
                                                            "Touchdown",
                                                            "Opp_Touchdown"))
```


Cleaning Distance.to.First.Down
```{r}
cmu.df$Distance.to.First.Down2 <- as.numeric(ifelse(cmu.df$Distance.to.First.Down == "Goal",
                                                cmu.df$Distance.to.Touchdown,
                                                cmu.df$Distance.to.First.Down))
cmu.df$is.Goal.Down <- ifelse(cmu.df$Distance.to.First.Down == "Goal",
                          1, 0)
```

Add a game.ID column and a fold column
```{r}
cmu.df$game.ID <- paste(cmu.df$Date, cmu.df$Home.Team.Abbreviation, "x", cmu.df$Away.Team.Abbreviation, sep="")

# fold
set.seed(123)
rand_vals <- sample(1:5, 7787, replace=TRUE)
cmu.df$fold <- rand_vals
```

```{r}
cmu.df <- subset(cmu.df, str_detect(cmu.df$game.ID, "CMU"))
```

Multinomial logistic regression for next score using down and distance (yds) to touchdown
```{r}
cmu.ep.model <- multinom(factor(Next_Score_Half) ~ 
                             factor(Down)
                           + log(Distance.to.First.Down2)
                           + Distance.to.Touchdown
                           + factor(is.Goal.Down)
                           + factor(Down):log(Distance.to.First.Down2), 
                           data = cmu.df, maxit = 300)
summary(cmu.ep.model)
```


Expected Points Column
```{r}
# get base probabilities
cmu_base_ep_preds <- as.data.frame(predict(cmu.ep.model, newdata = cmu.df, type = "probs"))
cmu_base_ep_preds <- rename(cmu_base_ep_preds,
                        No_Score_Prob=No_Score,
                        Field_Goal_Prob=Field_Goal,
                        Opp_Field_Goal_Prob=Opp_Field_Goal,
                        Touchdown_Prob=Touchdown,
                        Opp_Touchdown_Prob=Opp_Touchdown)
```

```{r}
# add to df
cmu.df <- cbind(cmu.df, cmu_base_ep_preds)
```

```{r}
cmu.EP <- rep(0,1406)
# multiply by point values
for (ii in 1:nrow(cmu.df)) {
  cmu.EP[[ii]] = sum(c(cmu.df[ii,38]*0, cmu.df[ii,39]*3, cmu.df[ii,40]*-3, cmu.df[ii,41]*7, cmu.df[ii,42]*-7))
}
cmu.df$EP <- cmu.EP
```


Expected Points Added Column
```{r}
cmu.df <- cmu.df %>%
  group_by(game.ID) %>%
  mutate(EPA =
           ifelse(Team.Possession == lead(Team.Possession),
                  lead(EP) - EP,
                  -lead(EP) - EP)) %>%
  ungroup()
```


Save the EP/EPA columns to the df
```{r}
write.csv(cmu.df, "CMU-2022-cleaned-pbp-wEPA.csv")
```












