---
title: "Model-ScrapedData"
author: "Jordan Gilbert, Marion Haney"
date: '2023-03-21'
output:
  html_document:
    df_print: paged
---

```{r}
# Libraries
require(ggplot2)
require(tidyverse)
require(dplyr)
```

```{r}
# Read in data
data = read.csv("PAC-2022-pbp2.csv")
```

Helper function that takes in the current play index, a vector of the scoring play indices, and play-by-play data and returns the score type and drive number for the next score
```{r}
FindNextScore = function(play_index, scoring_plays, data) {
  # next score index
  next_score_index = scoring_plays[which(scoring_plays >= play_index)[1]]
  
  # if next score index has no more scores after the current play
  if (is.na(next_score_index) | (data$Quarter[play_index] %in% c("1st", "2nd") & data$Quarter[next_score_index] %in% c("3rd", "4th")) | (data$Quarter[play_index] %in% c("3rd", "4th") & data$Quarter[next_score_index] %in% c("1st", "2nd", "OT"))) {
    score_type = "No Score"
    score_drive = data$Drive.Number[play_index]
  } 
  # else return the observed next score type and drive number
  else {
    # score drive number
    score_drive = data$Drive.Number[next_score_index]
    
    # TD
    if (data$Points.Scored.[next_score_index] & data$Touchdown.[next_score_index]) {
      # check who scored touchdown
      # look at extra point to know who scored
      if (!identical(data$Team.Possession[play_index], data$Team.Possession[next_score_index + 1])) {
        score_type = "Opp_Touchdown"
      }
      else {
        score_type = "Touchdown"
      }
    }
    # FG
    else if (data$Points.Scored.[next_score_index] & data$Field.Goal.[next_score_index]) {
      # check who scored field goal
      if (!identical(data$Team.Possession[play_index], data$Team.Possession[next_score_index])) {
        score_type = "Opp_Field_Goal"
      }
      else{
        score_type = "Field_Goal"
      }
    }
    # Extra Point
    else if (data$Points.Scored.[next_score_index] & data$Extra.Point.[next_score_index]) {
      # check who scored extra point
      if (!identical(data$Team.Possession[play_index], data$Team.Possession[next_score_index])) {
        score_type = "Opp_Extra_Point"
      }
      else {
        score_type = "Extra_Point"
      }
    }
    # Safety
    else if (data$Points.Scored.[next_score_index] & data$Safety.[next_score_index]) {
      # check who scored safety
      if (!identical(data$Team.Possession[play_index], data$Team.Possession[next_score_index + 1])) {
        score_type = "Opp_Safety"
      }
      else {
        score_type = "Safety"
      }
    }
    # Error
    else {
      score_type = NA
    }
  }
  return(data.frame(Next_Score_Half = score_type,
                    Drive_Score_Half = score_drive))
}
```


```{r}
# Which rows are the scoring plays
scoring_plays = which(data$Points.Scored.)
```


```{r}
next_score = lapply(c(1:nrow(data)), FindNextScore, scoring_plays = scoring_plays, data = data) %>% bind_rows()
data = cbind(data, next_score)
```

Drive Outcome
```{r}
GetDriveOutcome = function() {
  last_play_of_drive = which(data$Team.Possession != lag(data$Team.Possession))
  last_play_of_drive = last_play_of_drive - 1
  #last_play_of_drive_type = data$Outcome
  index = 1
  curr_last_play_index = last_play_of_drive[index]
  drive_outcome = c()
  for (curr_play in 1:length(last_play_of_drive)) {
    if (curr_play_index <= curr_last_play_index) {
      if (data$Touchdown.[curr_last_play_index]) {
        drive_outcome = append(drive_outcome, "")
      }
    }
  }
}
```


Quarters to Half
```{r}
## take away that weird NA row in the beginning
data <- slice(data, -1)
data$Half <- with(data, ifelse(Quarter %in% c("1st", "2nd"), "1st", "2nd"))
```


Save the cleaned data
```{r}
write.csv(data, "PAC-2022-Cleaned-pbp.csv", row.names=FALSE)
```



MODELING

Load in data & libraries
```{r}
require(nnet)
require(caret)
df <- read.csv("PAC-2022-Cleaned-pbp.csv")
```


Multinomial logistic regression for next score using down and distance (yds) to touchdown
```{r}
# Y: df$Next_Score_Half
# X: df$Down, log(df$Distance.to.Touchdown)
mult.log.model <- multinom(factor(Next_Score_Half) ~ factor(Down)
                           + log(Distance.to.Touchdown), data = df, maxit = 300)
summary(mult.log.model)
### how do i change the reference category to "No Score"?
```

Model Visualizations:
From: 
https://www.stat.cmu.edu/cmsac/sure/2022/materials/lectures/slides/21-Multinomial-multilevel.html#20 
```{r}
init_loso_cv_preds <- map_dfr(unique(df), 
          function(x) {
            # Separate test and training data:
            ## 75% of the sample size
            smp_size <- floor(0.75 * nrow(df))
            set.seed(1)
            train_ind <- sample(seq_len(nrow(df)), size = smp_size)
            train_data <- df[train_ind, ]
            test_data <- df[-train_ind, ]
            
            # Fit multinomial logistic regression model:
            ep_model <- multinom(factor(Next_Score_Half) ~ factor(Down)
                           + log(Distance.to.Touchdown), data = train_data, maxit = 300)
            
            # Return dataset of class probabilities:
            predict(ep_model, newdata = test_data, type = "probs") %>%
              as_tibble() %>%
              mutate(Next_Score_Half = test_data$Next_Score_Half)
            })
```

Calibration results for each scoring event
```{r}
ep_cv_loso_calibration_results <- init_loso_cv_preds %>%
  pivot_longer(Extra_Point:Opp_Field_Goal,
               names_to = "next_score_type",
               values_to = "pred_prob") %>%
  mutate(bin_pred_prob = round(pred_prob / 0.05) * .05) %>%
  group_by(next_score_type, bin_pred_prob) %>%
  summarize(n_plays = n(), 
            n_scoring_event = length(which(Next_Score_Half == next_score_type)),
            bin_actual_prob = n_scoring_event / n_plays,
            bin_se = sqrt((bin_actual_prob * (1 - bin_actual_prob)) / n_plays)) %>%
  ungroup() %>%
  mutate(bin_upper = pmin(bin_actual_prob + 2 * bin_se, 1),
         bin_lower = pmax(bin_actual_prob - 2 * bin_se, 0))
```

```{r}
ep_cv_loso_calibration_results <- na.omit(ep_cv_loso_calibration_results)
ep_cv_loso_calibration_results %>%
  mutate(next_score_type = fct_relevel(next_score_type, "Extra_Point", "Field_Goal", 
                                       "Opp_Extra_Point", "No Score", "Opp_Field_Goal"),
  next_score_type = fct_recode(next_score_type,
                               "Field Goal (3)" = "Field_Goal", 
                               "Touchdown (7)" = "Extra_Point",
                               "No Score (0)" = "No Score",
                               "-Field Goal (-3)" = "Opp_Field_Goal",
                               "-Touchdown (-7)" = "Opp_Extra_Point")) %>%
  ggplot(aes(x = bin_pred_prob, y = bin_actual_prob)) +
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = n_plays)) +
  geom_errorbar(aes(ymin = bin_lower, ymax = bin_upper)) + #coord_equal() +   
  scale_x_continuous(limits = c(0,1)) + 
  scale_y_continuous(limits = c(0,1)) + 
  labs(size = "Number of plays", x = "Estimated next score probability", 
       y = "Observed next score probability") + 
  theme_bw() + 
  theme(strip.background = element_blank(), 
        axis.text.x = element_text(angle = 90), 
        legend.position = c(1, .05), legend.justification = c(1, 0)) +
  facet_wrap(~ next_score_type, ncol = 4)
```





